---
title: "Coursera package - Luis Cano "
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=F,echo=TRUE}
library(coursera)
```

```{r,eval=F,echo=TRUE}
fars_read <- function(filename) {
        if(!file.exists(filename))
                stop("file '", filename, "' does not exist")
        data <- suppressMessages({
                readr::read_csv(filename, progress = FALSE)
        })
        dplyr::tbl_df(data)
}

dat <- fars_read("accident_2013.csv.bz2")

```

@ Function fars_read: import file to environemnt

@Arguments:
 filename : file to import

@ Value:
# return a dataframe 

@ Condition to error
 If file name doesn't exist

@ Require
 Function read_csv function in readr package and tbl_df function in and dplyr package  

@ examples
 data <- fars_read(file)
 class(data) 
 [1] "tbl_df"     "tbl"        "data.frame"

```{r,eval=F,echo=TRUE}
make_filename <- function(year) {
        year <- as.integer(year)
        sprintf("accident_%d.csv.bz2", year)
}

```

@# Function make_filename : assign year inside of the file name.
@ Arguments: 
@ year: year to add in the name file 

@ Value
 return a name of data with the value oif year inside it

@ Condition to error
If value of year is string  

@ Require
 sprintf function in r base package

@ Examples
 make_filename(1920)
 [1] "accident_1920.csv.bz2"

```{r,eval=F,echo=TRUE}
fars_read_years <- function(years) {
        lapply(years, function(year) {
                file <- make_filename(year)
                tryCatch({
                        dat <- fars_read(file)
                        dplyr::mutate(dat, year = year) %>% 
                                dplyr::select(MONTH, year)
                }, error = function(e) {
                        warning("invalid year: ", year)
                        return(NULL)
                })
        })
}

```

@ Function  "fars_read_years" : use differents years for create a diferent data frame with the values of select year and all month inside. 
 Arguments:
 years : vector of integrer or double with years

@ Value
 data frame with the values of select year and all month inside.

@ Condition to error
 If value of years doesn´t exist ini orinigal data frame , function doesn´t work

@ Require
 function mutate and select from dplyr package

@ Examples 
 fars_read_years(c(2018,2019,2020))

```{r,eval=F,echo=TRUE}
fars_summarize_years <- function(years) {
        dat_list <- fars_read_years(years)
        dplyr::bind_rows(dat_list) %>% 
                dplyr::group_by(year, MONTH) %>% 
                dplyr::summarize(n = n()) %>%
                tidyr::spread(year, n)
}

```

@ Function fars_summarize_years : take select years for create q new df with the summary of event in this periods
 Arguments:
 years : vector of integrer or double with years

@ Value
 data frame with the summary of event in this periods

@ Condition to error
 If value of years doesn´t exist ini orinigal data frame , function doesn´t work

@ Require
 group_by, summarize and bind_rows functions in dplyr package and spread function in tidyr package

@ Examples 
 fars_summarize_years(c(2019,2018,2020))

```{r,eval=F,echo=TRUE}
fars_map_state <- function(state.num, year) {
        filename <- make_filename(year)
        data <- fars_read(filename)
        state.num <- as.integer(state.num)

        if(!(state.num %in% unique(data$STATE)))
                stop("invalid STATE number: ", state.num)
        data.sub <- dplyr::filter(data, STATE == state.num)
        if(nrow(data.sub) == 0L) {
                message("no accidents to plot")
                return(invisible(NULL))
        }
        is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
        is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
        with(data.sub, {
                maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
                          xlim = range(LONGITUD, na.rm = TRUE))
                graphics::points(LONGITUD, LATITUDE, pch = 46)
        })
}
``` 

@ Function fars_map_state : create a map plot of the state according to the year selected
 Arguments:
 state : number of state 
 year : year to select 

@ Value
 plot (maps) with coordinates of state an number of cases

@ Condition to error
 Incorrect year in data, incorrect state number

@ Require
 filter function in dplyr package, maps function in maps package and points function in graphics package 

@ Examples
 fars_map_state(state.num=1, year = 2013)

